Analysis

### 2019-08-17 reports/manuscript.Rmd
finished introduction

One possible approach to this problem could be through formation of categories. The semantic categories in our mind are formed to match the categories of outside world and to be economic with terms of used properties to describe the structure [@rosch1978].

>>>>

One possible approach to this problem could be through formation of categories. The semantic categories in our mind are formed to match the categories of outside world and to be economic with terms of used properties to describe the structure [@rosch1978]. The category is formed around a prototype - a typical exemplar, which shows the maximal resemblance with rest of the members in the category, while having a low resemblance with other categories [@rosch1975b]. The prototypes are created not only for semantic categories, but also for visual stimuli. There are two main theories around visual prototype formation. The prototype-distance model [@Posner1970] states that the prototype is formed to minimize the distance from the presented stimuli, while the attribute-frequency model [@Neumann1974] states that prototype consist of attributes that were presented with highest frequency. Further studies showed higher experimental support for prototype-distance model for geometric patterns [@Franks1971] or faces [@Solso1979,@Solso1993,@Sampaio2018]. In the context of visual scenes, similar findings were not studied as it was difficult how to measure, what is a visual prototype.

Recent years brought possible solution to problem of visual prototypesimilarity of scenes. @Oliva2001 created a GIST descriptor that characterizes the spatial layout of the scene and it can be abstracted to natural properties of the scenes [@Torralba2003], which human observers can naturally use [@greene2009, @greene2010]. For this descriptor, we could compute pairwise distances, which would allows us to determine, which allows us to create an similarity space. With advances in training the convolutional neural networks [CNN, @Lecun2015,@goodfellow2016], they CNNs were widely used for various task in vision science in areas like scene categorization [@Krizhevsky2012], object recognition [@Kheradpisheh2016] or memorability of the image [@khosla2015]. One particular use of CNNs is to use the pretrained network, present it with a stimuli and use activation in one of the layers as a descriptor of the stimuli. The descriptors extracted in this way are knows as *deep features*. Again, we can compute distance between the deep features using L2 norm. @lukavsky2017 used this approach in a memory study to explore the properties of the similarity space defined by deep features.

In this study, we explore the properties of similarity space (which we denote as *image space*) defined by neural networks in two experiments. First, we explored whether the distances in the image space are understadable by human observers. Do humans understand, which scenes are closer to each other as defined by the image space? Second, we use the image space to verify, whether the visual prototypes are formed for complex scenes similarly as for simpler stimuli.
